---
title: "Seaweeds"
author: "Lilith Diener"
date: "2025-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

Firstly, we install and load all necessary packages

```{r}

# Install packages
# install.packages("ggmap")
# install.packages("ggmap", dependencies = TRUE)
# install.packages("ggspatial")
# install.packages("osmdata")
# install.packages("wesanderson") 

# Load packages
library(sf)
library(tidyverse)
library(ggplot2)
library(rinat)
library(rosm)
library(ggspatial)
library(ggmap)
library(osmdata)
library(leaflet)
library(htmltools)
library(mapview)
library(leafpop)
library(wesanderson)
library(dplyr)
```

Now we can read in our first data set - the South African Marine Protected Areas.
I am checking the data set's CRS and looking at the first few rows of the 
attribute table. 

```{r}
# Read in the Marine Protected Areas data set
mpa <- st_read("C:\\Users\\Win10\\Documents\\GIT\\Gracilaria\\Gracilaria\\data\\SAMPAZ_OR_2024_Q3.shp")
 
# Check the CRS
st_crs(mpa)

class(mpa)
head (mpa)

# Write data with sf
 st_write(mpa, "C:\\Users\\Win10\\Documents\\GIT\\Gracilaria\\Gracilaria\\data\\SAMPAZ_OR_2024_Q3_duplicate.shp", append = FALSE)
 
file.exists("C:\\Users\\Win10\\Documents\\GIT\\Gracilaria\\Gracilaria\\data\\SAMPAZ_OR_2024_Q3_duplicate.shp")
```

Here, I generated some plots to get an overview of the MPA data, I reduced it to
show only the column with MPA names and created an additional plot using tidyverse
that creates a legend. However, now the map is squashed.

```{r}
# Plot all (all columns, all of SA)
plot(mpa)
```

Figure 1:...

```{r}
# Plot only column 2 from attribute table, all of SA
 plot(mpa[2])
```

Figure 2: ...

```{r}
# Plot using tidyverse
ggplot() + geom_sf(data=mpa, aes(fill = `CUR_NME`))
```

Figure 3: ...

To avoid squashing the map, I crop it to the section I want to show. I set the 
bounding box to show the Cape Peninsula, and extend west and north to include 
Robben Island and east and south to include more of False Bay.

```{r}
 # Define bounding box size
xmin <- 18.0   # Extend west to include Robben Island
xmax <- 19.2   # Extend east to include parts of False Bay
ymin <- -34.6  # Extend south to cover more of False Bay
ymax <- -33.7  # Extend north to include Robben Island
```

I created a bounding box with the same CRS as my MPA data, cropped it to the 
Cape Peninsula region and surrounding areas, and plotted this. 

```{r}
# Create a bounding box with the same CRS as the MPA data
cape_bbox <- st_bbox(c(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), crs = st_crs(mpa))
 
# Crop the MPA data to the chosen region
mpa_cape <- st_crop(mpa, cape_bbox)

# Plot only cropped section of Cape Peninsula
plot(st_geometry(mpa_cape), col = 'lightblue', main = "Marine Protected Areas")
```

Figure 4: ...

I created a plot using ggplot2 showing the four MPAs included in my cropped 
section in different colours. For this to work, I first converted my MPA data
to a data frame. 

```{r}
# Convert MPA data to a data frame for ggplot2
mpa_cape$MPA_NAME <- as.factor(mpa_cape$CUR_NME)  # Convert MPAs to factor values

# Plot colour-coded MPAs
ggplot() +
  geom_sf(data = mpa_cape, aes(fill = CUR_NME), color = "black") +  
  scale_fill_viridis_d(name = "MPA Name") +  
  ggtitle("Marine Protected Areas - Cape Peninsula") +
  theme_minimal() +
  theme(legend.position = "right")  
```

Figure 5: ...

There are some adjacent polygons that belong to the same MPA. Below I edited 
the plot to not show any unwanted boundaries. 

```{r}
# Remove boundaries within MPAs
mpa_cape %>% group_by(CUR_NME) %>% 
  summarize() %>% 
  ggplot() + geom_sf(aes(fill = CUR_NME)) +
  scale_fill_discrete(name = "Marine Protected Areas") 
```

Figure 6: ...

Now it is time to add my species occurrence data from iNaturalist.
I am using three species: the kelps *Ecklonia maxima* and *Laminaria pallida*,
and the red algae *Gelidium pristoides*. I do this using the rinat() package.

```{r}
# Call species occurrence data from iNaturalist

# Ecklonia maxima
em <- get_inat_obs(taxon_name = "Ecklonia maxima",
                   bounds = c(-34.6, 18.0, -33.7, 19.2),
                   maxresults = 1000)

# View the first few rows of data
head(em)

# Gelidium pristoides
gpr <- get_inat_obs(taxon_name = "Gelidium pristoides",
                   bounds = c(-34.6, 18.0, -33.7, 19.2),
                   maxresults = 1000)

#View the first few rows of data
head(gpr)

# Laminaria pallida
lp <- get_inat_obs(taxon_name = "Laminaria pallida",
                   bounds = c(-34.6, 18.0, -33.7, 19.2),
                   maxresults = 1000)

# View the first few rows of data
head(lp)
```

I now merge the 3 datasets into a single data frame and filter the data so that
it only shows research grade observations and records from the wild (not found
in aquaculture farming systems). 

```{r}
# Merge the 3 datasets into a single data frame 
algae_data <- rbind(gpr, em, lp)

# Check structure
str(algae_data)

# Filter returned observations by a range of column attribute criteria
algae_data <- algae_data %>% filter(positional_accuracy<46 & 
                latitude<0 &
                !is.na(latitude) &
                captive_cultivated == "false" &
                quality_grade == "research")

class(algae_data) # check class
```

I convert the data frame into a spatial object. I define the CRS to be to be Geographic WGS84 using the EPSG code.

```{r}
# Make the data frame a spatial object of class = "sf"
algae_data <- st_as_sf(algae_data, coords = c("longitude", "latitude"), crs = 4326)

class(algae_data) # confirm new classes

# See the new "geometry" column
names(algae_data)
```

I created my first plot that includes the 3 species occurrences within the MPAs.

```{r}
# Plot 3 seaweed species with MPAs and legend
ggplot() +
  geom_sf(data = mpa_cape, aes(fill = CUR_NME), color = "black", alpha = 0.5) +
  geom_sf(data = algae_data, aes(color = scientific_name), size = 2) +
  scale_fill_viridis_d(name = "Marine Protected Areas") +  
  scale_color_manual(name = "Seaweed Species", values = c("red", "green", "blue", "purple")) + 
  coord_sf(xlim = c(18.0, 19.2), ylim = c(-34.6, -33.7)) +
  ggtitle("MPAs and Algae Observations") +
  theme_minimal() +
  theme(legend.position = "right")  
```

Figure 7: ...

I created an interactive map with a default OSM base map that allows me to zoom 
into specific areas and has popup labels with URL links to each observation
on iNaturalist.

```{r}
# Make an interactive map
leaflet() %>%
  addTiles(group = "Default") %>%  
  addCircleMarkers(data = algae_data,
                   group = "Seaweeds",
                   radius = 3, 
                   color = "green") 
```

Figure 8: ...

I am adding popup labels that show the Feature ID, the particular species' 
scientific name, and whether they are cultivated.

```{r}
# Add popup labels 
mapview(algae_data, 
        popup = 
          popupTable(algae_data,
            zcol = c("scientific_name", "captive_cultivated", "url")))
```

Figure 9: ...

Now I am adding URL links to each observation on iNaturalist. 

```{r}
# Make URLs links
l_algae_data <- algae_data %>%
  mutate(click_url = paste("<b><a href='", url, "'>Link to iNat observation</a></b>"))

mapview(algae_data, 
        popup = 
          popupTable(l_algae_data,
            zcol = c("scientific_name", "captive_cultivated", "click_url")))
```

I want to now intersect the MPA and species data sets. The CRS of the two do not match, so I first transform one data set to match the other and then intersect.

```{r}
# Transform one data set to match the other
algae_data <- st_transform(algae_data, st_crs(mpa_cape))

# Confirm CRS of both data sets
st_crs(algae_data)
st_crs(mpa_cape)

# Intersect species occurrences and MPAs
mpa_em_lp_gpr <- st_intersection(algae_data, mpa_cape)

# Check dimensions before and after intersection
dim(algae_data) # before
algae_data <- st_intersection(algae_data, mpa_cape) # intersect
dim(algae_data) # after
```

Now I create a plot of the intersected data with an OSM base map. The species
records from different MPAs are indicated in different colours.

```{r}
# WEGLASSEN?
# Plot intersected dataset
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=algae_data)

# Colour points
pal <- wes_palette("Darjeeling1", 7, type = "continuous")

ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = algae_data, aes(col = CUR_NME)) +
  scale_colour_manual(name = "Marine Protected Areas", values = pal)
```

Figure 10: ...

It looks like almost all species records are in the Table Mountain National Park Marine Protected Area. Let us confirm this.

```{r}
algae_data %>% group_by(CUR_NME) %>% summarise(n())
```

When we have a look at the numbers in column n(), this proves to be correct.
It is difficult to see the few records within the Betty's Bay (BB) MPA and Robben Island (RI) MPA are on the map. I am therefore labelling these two.

```{r}
# Find BB MPA localities
bb <- algae_data %>% filter(CUR_NME %in% c("Betty's Bay Marine Protected Area")) 

# Find RI MPA localities
ri <- algae_data %>% filter(CUR_NME %in% c("Robben Island Marine Protected Area"))  

# Label Betty's Bay (BB) MPA and Robben Island (RI) MPA 
ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data = algae_data, aes(col = CUR_NME)) +
  scale_colour_manual(values = pal) +
  labs(color = "Marine Protected Areas") +  # Rename the legend title here
  geom_sf_label(data = bb %>% filter(CUR_NME == "Betty's Bay Marine Protected Area"), aes(label = "BB")) +
  geom_sf_label(data = ri %>% filter(CUR_NME == "Robben Island Marine Protected Area"), aes(label = "RI")) +
  theme_minimal()
```

Figure 11: ...

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
